Discriminating the Influence of Correlated Factors \\from Multivariate Observations: the Back-to-Back Regression

### 2020-05-19 paper/fmri_exps.tex
up

Next, we apply our method to brain imaging data from the anonymized multimodal
neuroimaging ``Mother Of all Unification Studies'' (MOUS) dataset
\citep{schoffelen2019204}. The dataset contains functional Magnetic Resonance
Imaging (fMRI) and magneto-encephalography (MEG)
recordings of each 102 healthy native-Dutch adults who participated in a reading
task. Thirteen subjects were excluded from the analysis (12 MEG and 1 fMRI)
because of missing files or corrupted headers.
%
Subjects were exposed to a rapid serial visual presentation of Dutch words. The
word lists consisted of 120 sentences, and scrambled lists of the same words.
Each word was presented on the computer screen for 351ms on average (min: 300ms,
max: 1400ms). Successive words were separated by a blank screen for 300ms, and
successive sentences were separated by an empty screen for a few (3-4) seconds.

\subsubsection{fMRI preprocessing}
Results included in this manuscript come from preprocessing performed
using \emph{fMRIPrep} 20.0.7 (\citet{fmriprep1}; \citet{fmriprep2};
RRID:SCR\_016216), which is based on \emph{Nipype} 1.4.2
(\citet{nipype1}; \citet{nipype2}; RRID:SCR\_002502).

@@

A total of 2 T1-weighted (T1w) images were found within the input BIDS

@@

Asymmetrical template version 2009c} {[}\citet{mni152nlin2009casym},

@@

Stereotaxic Registration Model} {[}\citet{mni152nlin6asym},

@@

For each of the 2 BOLD runs found per subject (across all tasks and
sessions), the following preprocessing was performed. First, a reference

@@

\subsubsection{fMRI analysis}

The preprocessed volumetric fMRI data was linearly projected to its closest surface
using nilearn $vol_to_surf$ function with a 3 mm radius. A surface search-light
analysis was then implemented by concatenating the surface
fMRI data within a 8mm radius of each vertex. For each subject and vertex, we
thus build an observation matrix $Y \in \mathbb{R}^{n \times d_y}$ of $n\approx$ 1400 words
by $d_y\approx40$ vertices. Each
of the columns of $Y$ is normalized to have zero mean and unit variance.

@@

We aim to identify the word features that cause a variation in brain responses. We consider four distinct but colinear features.
%
First, 'Word Length' refers to the total number of letters. Word Length is expected to specifically cause a variation in the early evoked MEG responses (i.e. from 100 ms after stimulus onset) elicited by the retinotopically-tuned visual cortices (e.g. \citep{pegado2014timing}.).
%
Second, 'Word Frequency' indexes how frequently each word appears in Dutch and was derived with the the Zipf logarithmic scale of \citep{van2014subtlex} provided by the WordFreq package \citep{speerwordfreq}. Word Frequency is expected to specifically cause a variation in the late evoked MEG responses (i.e. from 400 ms), because it variably engages semantic processes in the temporal cortices \citep{kutas2011thirty}.
%
Third, 'Word Function' indicates whether each word is a content word (i.e. a noun, a verb, an adjective or an adverb) or a function word (i.e. a preposition, a conjunction, a determinant, a pronoun or a numeral), and was derived from Spacy's part of speech tagger \citep{spacy2}. To our knowledge, this feature has not been thouroughly investigated with MEG. Its causal contribution to reading processes in the brain thus remains unclear.
%
Finally, to verify that B2B and other methods would not inadequately identify non-causal features, we added a dummy feature, constructed from a noisy combination of Word Length and Word Frequency:
$dummy = z(length) + z(frequency) + \mathcal{N}$, where $z$ normalizes features and $\mathcal{N}$ is a random vector sampling Gaussian distribution (all terms thus have a zero-mean and a unit-variance). This procedure yields an $X \in \mathbb{R}^{n \times d_x}$ matrix of $n\approx$ 2700 words by
$d_x=4$ features for each subject. Each of the columns of $X$ is normalized to
have a mean and a standard deviation of 0 and 1 respectively.

To account for the delay of blood oxygenation level dependent responses, the
features were convolved using the default hemodynamic response function Glover
of nilearn \citep{nilearn}.

@@

\begin{figure}
  \begin{center}
          \includegraphics[width=.8\textwidth,
                       trim=1cm 1cm 2cm 2cm,
                       clip=True]{figures/fmri_delta_r.png}
          \includegraphics[width=.8\textwidth,
                       trim=0cm 0cm 0cm 0cm,
                       clip=True]{figures/fmri_strip.png}

@@

\subsubsection{Results}
We compared the ability of Forward regression, Backward regression, CCA, PLS
and B2B to estimate the causal contribution of four distinct but collinear
features on brain responses to words.

Supplementary Figure \ref{fig:fmri_supp} shows that the Backward model decodes
the dummy variable well above chance level. In addition, it decodes both Word
Length and Word Frequency across large sections of the cortex, even though
these features are known to specifically influence early visual cortex and
associative cortices respectively. These results illustrate that backward
modeling cannot be used to estimate the causal contribution of collinear
features.

On the contrary, the $\beta$ coefficients of the Forward model reveals the
expected spatial specificity of Word Length and Word Frequency. However, these
coefficients are moderately consistent across subjects, and can thus
dramatically limit the possibility to achieve high levels of
significance, as evaluated with a Wilcoxon tests across subjects at each vertex.

In principle, the inter-subject variability could be less impactful with
multivariate observations methods such as CCA and PLS. However, these
cross-decompositions do not provide a single and interpretable coefficient
for each feature at each vertex.

On the contrary, the $\hat E$ coefficients of an unbiased B2B (i.e. a B2B where
$H$ is not regularized) derive from multivariate fMRI observations but are
unidimensional for each feature. In fact, $\hat E$ reaches higher levels of
significance, as evaluated with a Wilcoxon test across subjects at each vertex
than the Forward $\beta$ \ref{fig:fmri_supp}. Note that a direct
comparison between the Forward coefficients $\beta$ and those of B2B ($\hat E$)
is not feasible, as these two do not have the same units.

To fairly compare the Forward, CCA, PLS and B2B methods on a common metric,
without being over-dependent on the
potentially invalid assumptions of each model, we compare them on their ability
to improve Y
predictions of out-of-sample data when one factor is introduced into the model.
This analysis leads to a $\Delta R$ as described in Algorithm
\ref{algorithm:b2b_fi}). As expected, the Forward, CCA, PLS
and B2B method predicted that the Dummy Variable does not improve the $Y$
prediction.

The average $\Delta R$ across subjects is displayed for each method and each
feature in \ref{fig:fmri_r}, and suggest that B2B surpasses other methods.
To formally quantify these comparisons, we compare, for each subject separately,
the average $\Delta R$ across vertices. To avoid including irrelevant brain
regions, we restrict this analysis to vertices which could be reliably accounted
for by the Forward model ($p<.001$, not corrected across vertices)
\ref{fig:fmri_r}. The results show that B2B significantly outperforms all
other methods on all features (all $p<0.0001$) but the dummy variable
(min $p=0.0852$).

>>>>

Next, we apply our method to brain imaging data from the anonymized multimodal
neuroimaging ``Mother Of all Unification Studies'' (MOUS) dataset
\citep{schoffelen2019204}. The dataset contains functional Magnetic Resonance
Imaging (fMRI) and magneto-encephalography (MEG)
recordings of each 102 healthy Dutch adults who performed a reading
task in the scanner. Thirteen subjects were excluded from the analysis (12 MEG and 1 fMRI)
because of technical difficulties reading the files.
%
Subjects were exposed to a rapid serial visual presentation of Dutch words. The
word lists consisted of 120 sentences, and scrambled lists of the same words.
Each word was presented on the computer screen for 351ms on average (min: 300ms,
max: 1400ms). Successive words were separated by a blank screen for 300ms, and
successive sentences were separated by an empty screen for a few (3-4) seconds.

\subsubsection{fMRI preprocessing}
Results included in this manuscript come from preprocessing performed
using \emph{fMRIPrep} 20.0.7 (\citep{fmriprep1}; \citep{fmriprep2};
RRID:SCR\_016216), which is based on \emph{Nipype} 1.4.2
(\citep{nipype1}; \citep{nipype2}; RRID:SCR\_002502).

@@

A total of two T1-weighted (T1w) images per subject were found within the input BIDS

@@

Asymmetrical template version 2009c} {[}\citep{mni152nlin2009casym},

@@

Stereotaxic Registration Model} {[}\citep{mni152nlin6asym},

@@

For each of the 2 BOLD runs found per subject, the following preprocessing was performed. First, a reference

@@

The preprocessed volumetric fMRI data was linearly projected to its closest surface
using nilearn \texttt{vol\_to\_surf} function with a 3 mm radius. A surface searchlight
analysis was then implemented by concatenating the surface
fMRI data within a 8mm radius of each vertex. For each subject and vertex, we
thus build an observation matrix $Y \in \mathbb{R}^{n \times d_y}$ of $n\approx$ 1400 words
by $d_y\approx40$ vertices per searchlight sphere. Each
of the columns of $Y$ is normalized to have zero mean and unit variance.

@@

We aim to identify the word features that cause a variation in brain responses. We consider four distinct but colinear features.
%
First, 'Word Length' refers to the total number of letters. Word Length is expected to primarily cause a variation in the early evoked MEG responses (i.e. from 100 ms after stimulus onset) elicited by the retinotopically-tuned visual cortices (e.g. \citep{pegado2014timing}.).
%
Second, 'Word Frequency' indexes how frequently each word appears in Dutch and
was derived with the the Zipf logarithmic scale of \citep{van2014subtlex}
provided by the WordFreq package \citep{speerwordfreq}. Word Frequency is
expected to primarily cause a variation in the late evoked MEG responses (i.e.
from 400 ms) in the left frontal, temporal and parietal cortices
cortices \citep{kutas2011thirty,mccandliss2003visual}.
%
Third, 'Word Function' indicates whether each word is a content word (i.e. a
noun, a verb, an adjective or an adverb) or a function word (i.e. a preposition,
a conjunction, a determinant, a pronoun or a numeral), and was derived from
Spacy's part of speech tagger \citep{spacy2}. To our knowledge, this feature has
not been thoroughly investigated with fMRI and MEG. While its causal contribution to
reading processes in the brain thus remains unclear, this lexical feature can
nonetheless be expected to present similar brain patterns to word frequency.
%
Finally, to verify that B2B and other methods would not inadequately identify non-causal features, we added a dummy feature, constructed from a noisy combination of Word Length and Word Frequency:
$dummy = z(length) + z(frequency) + \mathcal{N}$, where $z$ normalizes features
and $\mathcal{N}$ is a random vector sampling Gaussian distribution (all terms
thus have a zero-mean and a unit-variance).

To account for the delay of blood oxygenation level dependent responses, the
features were convolved using the Glover hemodynamic response function
of nilearn's \texttt{compute\_regressor} function an oversampling of 16 and default parameters \citep{nilearn}.

This procedure yields an $X \in \mathbb{R}^{n \times d_x}$ matrix of $n\approx$
840 TR (Repetition Time: 2 sec.) by
$d_x=4$ features for each subject. Each of the columns of $X$ is normalized to
have a zero-mean and a unit variance.

@@

\begin{figure}
  \begin{center}
          \includegraphics[width=.8\textwidth,
                       trim=1cm 1cm 1cm 2cm,
                       clip=True]{figures/fmri_delta_r.png}
          \includegraphics[width=.8\textwidth,
                       trim=0cm 0cm 0cm 0cm,
                       clip=True]{figures/fmri_strip.png}

@@

\subsubsection{fMRI Results}
We compared the ability of Forward regression, Backward regression, CCA, PLS
and B2B to estimate the causal contribution of four distinct but collinear
features on brain responses to words.

Supplementary Figure \ref{fig:fmri_supp} shows that the Backward model decodes
the dummy variable well above chance level. In addition, it decodes both Word
Length and Word Frequency across large sections of the cortex, even though
these features are known to primarily influence early visual cortex and
associative cortices respectively. These results illustrate that while Backward
modeling achieve high sensitivity, it is not valid to estimate the specific
causal contribution of collinear features.

On the contrary, the $\beta$
coefficients of the Forward model reveals the expected spatial specificity of
Word Length and Word Frequency, peaking the visual and temporal cortices
respectively. However such Forward coefficients may be variable across
subjects, and could thus underestimate the significance of each factor at each
vertex, as evaluated with a Wilcoxon tests across subjects. In principle, such
inter-subject variability could be less impactful with
multivariate observations methods such as CCA and PLS \citep{bilenko2016pyrcca,
king2018encoding}, but they fail to provide a single and interpretable coefficient
for each feature at each vertex.

In contrast, the $\hat E$ of an unbiased B2B (i.e. a B2B where $H$ is not
regularized) reaches higher levels of significance at each vertex
than the Forward $\beta$ \ref{fig:fmri_supp}. (Note that a direct
comparison between the Forward coefficients $\beta$ and those of B2B ($\hat E$)
is not feasible, as these two do not have the same units.)

To fairly compare the Forward, CCA, PLS and B2B methods on a common and robust
evaluation metric we compare them on their ability to improve $Y$
predictions of out-of-sample data when one factor is introduced into the model.
This analysis leads to a $\Delta R$ as described in Algorithm
\ref{algorithm:b2b_fi}). As expected, the Forward, CCA, PLS
and B2B method predicted that the Dummy Variable does not improve the $Y$
prediction. This confirms that these methods accurately rule out the known
non-causal factor.

The average $\Delta R$ across subjects is displayed for each model and each
feature in \ref{fig:fmri_delta_r}. Overall, B2B favorably compares to baseline models.
To quantify this assessment, we compare for each subject separately
the average $\Delta R$ across vertices obtained between B2B and each
baseline model. To limit the inclusion of uninformative brain regions in this
summary, we restrict the analysis to vertices which can be reliably accounted
for by the Forward model ($p<.001$, not corrected across vertices)
\ref{fig:fmri_delta_r}. The results show that B2B significantly outperforms all
baseline models on all factors (all $p<0.0001$) but the dummy variable
(min $p=0.0852$).

### 2020-05-19 paper/meg_exps.tex
up

As expected, the Backward model reveals a similar decoding time course for Word
Length and Word Frequency, even though these features are known to specifically
influence early and late MEG responses respectively \citep{kutas2011thirty}. In
addition, the same decoding time course was observed for the dummy variable.
These results illustrate that backward modeling cannot be used to estimate the
causal contribution of correlated features.

@@

Word Length and Word Frequency improved the prediction performance of all
methods: $\Delta R>0$ for all models (all $p<0.0001$). As expected, the time
course associated with Word Length and Word Frequency rose from $\approx$ 100 ms
and from $\approx$ 400 ms respectively. Furthermore, Word Function improved the
prediction performance of all models (all $p < 0.0002$) except for
PLS~($p=0.7989$). Overall, these results confirm that Word Length, Word
Frequency and Word Function causally influence specific periods of brain
responses to words.

To assess which model would be most sensitive to these causal discoveries, we
compared B2B to other models across subjects (Figure \ref{fig:meg_results}
right). For Word Length B2B outperforms all models (all $p < 0.00001$) but CCA
($p=0.0678$). For Word Frequency, B2B outperforms all models (all $p < 0.0006$).
For "Word Function", B2B outperforms all models (all $p < 0.0015$). Overall,
these results show that B2B reliably outperforms standard methods, especially
when the effects are difficult to detect.

>>>>

As expected, the Backward model reveals a similar decoding time course for Word
Length and Word Frequency, even though these features are known to specifically
influence early and late MEG responses respectively \citep{kutas2011thirty}. In
addition, the same decoding time course was observed for the dummy variable.
Once again, these results illustrate that Backward modeling cannot be used to estimate the
causal contribution of correlated features.

@@

Word Length and Word Frequency improved the prediction performance of all
methods: $\Delta R>0$ for all models (all $p<0.0001$). As expected, the time
course associated with Word Length and Word Frequency rose from $\approx$ 100 ms
and from $\approx$ 400 ms respectively. Furthermore, Word Function improved the
prediction performance of all models (all $p < 0.0002$) except for
PLS~($p=0.7989$). Overall, these results confirm that Word Length, Word
Frequency and Word Function specifically improve the prediction of specific periods of brain
responses to words, and thus form plausible independent causal contributors.

To assess which model would be most sensitive to these causal discoveries, we
compared B2B to other models across subjects (Figure \ref{fig:meg_results}
right). For Word Length, B2B outperforms all models (all $p < 0.00001$) but CCA
($p=0.0678$). For Word Frequency, B2B outperforms all models (all $p < 0.0006$).
For "Word Function", B2B outperforms all models (all $p < 0.0015$). Overall,
these results show that B2B favorably compares against baseline models.

### 2020-05-19 paper/paper.tex
up

We consider the measurement of multivariate signal $Y \in \mathbb{R}^{n \times
d_y}$ (the dependent variables), generated from a set of putative causes $X \in \mathbb{R}^{n \times
d_x}$ (the independent variables), via some unknown linear apparatus $F \in \mathbb{R}^{d_x \times d_y}$.
%
Not all the variables in $X$ exert a causal influence on $Y$.
%
By considering a square binary diagonal matrix of \emph{causal influences} $E
\in \mathbb{D}^{d_x \times d_x}$, we denote by $XE$ the causal factors of $Y$.
%
In summary, the problem can be formalized as:
%
\begin{equation}
    y_i = (x_i E + n_i) F
    \label{eq:model}
\end{equation}
%
where $i$ is a given sample, and $n_i$ is a sample-specific noise drawn from a
centered distribution.
%
While the triplet of variables $X$ and $N$ are independent, we allow each of
them to have any form of covariance.
% what is a general covariance matrix?
%
In practice, we observe $n$ samples $(X, Y)$ from the model.
%
This problem space, along with the sizes of all variables involved, is
illustrated in Figure~\ref{fig:b2b}.
%
Given the model in Equation~\cref{eq:model}, the goal of Back-to-Back
Regression (B2B) is to estimate the matrix of $E$, i.e. to identify the factors
that most reliably account for the multivariate observations.

@@

Performing two regressions over the same data sample can result in overfitting,
as spurious correlations in the data absorbed by the first regression will be
leveraged by the second one.
%
To avoid this issue, we split our sample $(X, Y)$ into two splits $(X_1, Y_1)$
and $(X_2, Y_2)$.
%
Then, the first regression is performed using $(X_1, Y_1)$, and the second
regression is performed using $(X_2, Y_2)$.
%
To compensate for the reduction in sample size caused by the split, B2B is
repeated over many random splits, and the final estimate $\hat E$ of the causal
influence matrix is the average over the estimates associated to each split
\citep{breiman1996bagging}.
%
To accelerate this ensembling procedure, we implemented an efficient
leave-one-out cross-validation scheme as detailed in \citep{rifkin2007notes}
% http://cbcl.mit.edu/publications/ps/MIT-CSAIL-TR-2007-025.pdf,
as follows:
%

@@

In the presence of noise, the causal factors of $E$ correspond to the positive
elements of $\text{diag}(\hat H)$. The methods to recover them are presented in
the Appendix \ref{recovering}.

@@

\subsection{Magnetoencephalography}
\label{sec:experiment_real}
\input{meg_exps.tex}

@@

Forward and cross-decomposition models have been used to identify the causal
contribution of correlated factors onto multi-dimensional observations (e.g.
\citep{naselaris2011encoding}). These approaches typically lead to multiple
coefficients for each features (i.e. one per dimension of $Y$ or one per
component respectively). Furthermore, these coefficients can be difficult to
summarize into a single causal estimate. By contrast, B2B quickly (Fig.
\ref{fig:duration}) leads to a single unbiased scalar values $\hat E$ tending
towards 1 and 0 for causal and non-causal features respectively.

A variety of other statistical methods applied to neuroimaging data have been
proposed to clarify what is being represented in brain responses - i.e. what
feature causes specific brain activity. One of the popular linear method is
Representational Similarity Analysis (RSA)
\citep{kriegeskorte2008representational}, and consists in analyzing the
similarity of brain responses associated with specific categorical conditions
(e.g. distinct images), by (1) fitting one-against-all classifiers on each
condition and (2) testing whether these classifiers can discriminate all other
conditions. The resulting confusion matrix is then analyzed in an unsupervised
manner to reveal which conditions lead to similar brain activity patterns. B2B
differs from RSA in that (1) it uses regressions instead of classifications, and
can thus generalize to new items and new contexts and (2) it is fully
supervised.

% Here, we focus on the restricted issue of disentangling the
% influence of linearly correlated predictors ($X$) onto noisy multivariate
% observations ($Y$). The present approach thus differs from other causal
% discovery algorithms based on temporal-delays and/or nonlinear interactions in
% systems where the directionality of causation (from X to Y or vice versa) is
% unknown (e.g. \citep{peters2017elements, granger1969investigating,
% janzing2013quantifying, scholkopf2016modeling}.

Finally, CCA has been used in neuroimaging for a variety of purposes such as
denoising and subject alignment \citep{cca_hotelling, de2019multiway}. While CCA
relates to B2B, these two methods diverge in several ways. First, CCA and B2B
have different objectives: CCA aims to find the potentially numerous and poorly
interpretable components where X and Y are maximally correlated, whereas B2B
aims to recover the causal factors from X to Y. Second, B2B is not symmetric
between $X$ and $Y$: it aims to identify specific causal features by first
optimizing over the decoders $G$ and then over $H$. By contrast, CCA is
symmetric between $X$ and $Y$, and aims to find $G$ and $H$ such that they
project $X$ and $Y$ on maximally correlated dimensions. Third, CCA is based an
eigen decomposition of $XH$ and $YG$ - the corresponding canonical components
are thus mixing the $X$ features in way that limit interpretability and
potentially dilute the impact of each feature onto multiple components. In
contrast B2B assesses each feature $X^j$ on a single $Y$ component specifically
selected to maximize signal-to-noise ratio of that feature $j$. Fourth, and
unlike B2B, CCA does not separately optimize two distinct regularization
parameters for $G$ and $H$. Finally, CCA does not use different data splits to
estimate $G$ and $H$. Together, these differences may explain why B2B reliably
outperform CCA on estimating causal influences (Figs. \ref{fig:percondition} and
\ref{fig:auc_plots}).

% I find a bit surprising to mention causality literature in ML at the end. I
% would have expected this in the introduction to convince reviewer that know
% your stuff.

\section{Conclusion}
In this work, we proposed Back-to-Back (B2B) regression, a linear method to
disentangle confounded factors from multidimensional observations.
%
B2B repeatedly performs two successive multidimensional regressions on
independent subsets of the data: the first regression is applied on the output
domain (as in backward decoding), whereas the second regression is applied on
the input domain (as in forward encoding).
%
We provided a theoretical guarantee about the consistency of B2B, and compared
it to several baselines in controlled synthetic experiments.
%
We also applied B2B to a recent brain imaging dataset, analyzing the timing of
brain responses and their connection to word features.
%
We obtained results consistent with prior work in neuroscience literature,
confirming the reliability of B2B for real data analysis.
%

>>>>

\subsection{Problem setup}

We consider the measurement of multivariate signal $Y \in \mathbb{R}^{n \times
d_y}$ (the dependent variables), generated from a set of putative causes $X \in \mathbb{R}^{n \times
d_x}$ (the independent variables), via some unknown linear apparatus $F \in \mathbb{R}^{d_x \times d_y}$.
%
Not all the variables in $X$ exert a causal influence on $Y$.
%
By considering a square binary diagonal matrix of \emph{causal influences} $E
\in \mathbb{D}^{d_x \times d_x}$, we denote by $XE$ the causal factors of $Y$.
%
In summary, the problem can be formalized as:
%
\begin{equation}
    y_i = (x_i E + n_i) F
    \label{eq:model}
\end{equation}
%
where $i$ is a given sample, and $n_i$ is a sample-specific noise drawn from a
centered distribution.
%
While the triplet of variables $X$ and $N$ are independent, we allow each of
them to have any form of covariance.
% what is a general covariance matrix?
%
In practice, we observe $n$ samples $(X, Y)$ from the model.
%
This problem space, along with the sizes of all variables involved, is
illustrated in Figure~\ref{fig:b2b}.
%
Given the model in Equation~\cref{eq:model}, the goal of Back-to-Back
Regression (B2B) is to estimate the matrix $E$, i.e. to identify the factors
that most reliably account for the multivariate observations.

@@

Performing two regressions over the same data sample can result in overfitting,
as spurious correlations in the data absorbed by the first regression will be
leveraged by the second one.
%
To avoid this issue, we split our sample $(X, Y)$ into two splits $(X_1, Y_1)$
and $(X_2, Y_2)$.
%
The first regression is performed using $(X_1, Y_1)$, and the second
regression is performed using $(X_2, Y_2)$.
%
To compensate for the reduction in sample size caused by the split, the two
successive regressions are
repeated over many random splits, and the final estimate $\hat E$ of the causal
influence matrix is the average over the estimates associated to each split
\citep{breiman1996bagging}.
%
To accelerate this ensembling procedure, we use an efficient
leave-one-out cross-validation scheme as detailed in \citep{rifkin2007notes}
% http://cbcl.mit.edu/publications/ps/MIT-CSAIL-TR-2007-025.pdf,
as follows:
%

@@

In the presence of noise, the causal factors of $E$ correspond to the positive
elements of $\text{diag}(\hat H)$. The methods to recover them are presented in
the Appendix (\ref{recovering}).

@@

\subsection{Magneto-encephalography}
\label{sec:experiment_real}
\input{meg_exps.tex}

@@

% Forward and cross-decomposition models have been used to identify the causal
% contribution of correlated factors onto multi-dimensional observations (e.g.
% \citep{naselaris2011encoding}). These approaches typically lead to multiple
% coefficients for each features (i.e. one per dimension of $Y$ or one per
% component respectively). Furthermore, these coefficients can be difficult to
% summarize into a single causal estimate. By contrast, B2B quickly (Fig.
% \ref{fig:duration}) leads to a single unbiased scalar values $\hat E$ tending
% towards 1 and 0 for causal and non-causal features respectively.

A variety of statistical methods applied to neuroimaging have been
proposed to clarify what is being represented in brain responses - i.e. what
feature causes specific brain activity. For example, CCA and PLS have been used in
neuroimaging for a variety of purposes such as
denoising and subject alignment \citep{cca_hotelling, de2019multiway}. While CCA
and PLS relates to B2B, these methods diverge in several ways. First, they have different objectives: CCA aims to find the potentially numerous and poorly
interpretable components where $X$ and $Y$ are maximally correlated, whereas B2B
aims to recover the causal factors from X to Y. Second, B2B is not symmetric
between $X$ and $Y$: it aims to identify specific causal features by first
optimizing over the decoders $G$ and then over $H$. By contrast, CCA and PLS are
symmetric between $X$ and $Y$, and aims to find $G$ and $H$ such that they
project $X$ and $Y$ on maximally correlated dimensions. Third, CCA is based an
eigen decomposition of $XH$ and $YG$ - the corresponding canonical components
are thus mixing the $X$ features in way that limit interpretability and
potentially dilute the impact of each feature onto multiple components. In
contrast B2B assesses each feature $X^j$ on a single $Y$ component specifically
selected to maximize signal-to-noise ratio of that feature $j$. Fourth CCA does
not separately optimize two distinct regularization
parameters for $G$ and $H$, where B2B does. Finally, CCA does not use different data splits to
estimate $G$ and $H$. Together, these differences may explain why B2B reliably
outperform CCA on estimating causal influences (Figs. \ref{fig:percondition} and
\ref{fig:auc_plots}).

One popular multivariate analysis of brain activity is
Representational Similarity Analysis (RSA)
\citep{kriegeskorte2008representational}, and consists in analyzing the
similarity of brain responses associated with specific categorical conditions
(e.g. distinct images), by (1) fitting one-against-all classifiers on each
condition and (2) testing whether these classifiers can discriminate all other
conditions. The resulting confusion matrix is then analyzed, generally in an unsupervised
manner, to reveal which conditions lead to similar brain activity patterns. B2B
differs from RSA in that it uses regressions instead of one-hot classifications.
Consequently, and unlike RSA, B2B (1) is fully supervised (2) provides
interpretable coefficients and (3) can generalize to new items and new contexts.
In practice, these elements allow B2B to apply to event-related paradigms and
latent variable analyses.

The present empirical results, together with their theoretical foundations,
suggest that B2B may come as a tool of choice for causal discovery when
feature-wise intervention is limited.
